Multiscalar models for systems of cities

### 2023-10-12 Docs/Communications/2023/ECTQG2023/slides.tex

\sframe{Hybrid models}{

@@

@@

@@

% TODO histogram/distib for illustrated indicators for the innovation model?

@@

@@

\sframe{Discussion}{

% \textbf{Main results:}

@@

$\rightarrow$

@@

$\rightarrow$

@@

@@

>>>>

\sframe{Hybrid models mixing scales}{

@@

$\rightarrow$ models coupling scales with a single direction of inter-scale feedback can be simplified to non-stationary dynamics or fixed effects at a single scale: ``strong'' coupling is necessary

$\rightarrow$ scales have their own distinct ontologies with their own processes

% Q: is aggregation a process? not necessarily - but a feedback between variables

$\rightarrow$ inter-scales processes may be necessary to capture feedbacks and may necessitate their own ontology

@@

$\rightarrow$ model parametrisation with real data is more complicated: which scales? need of consistent datasets in case of multiple scales (cf Geodivercity project); how to consistently setup model states? \ldots

$\rightarrow$ synthetic systems of cities are a first way to explore model dynamics independently of geographical contingencies \cite{raimbault2019space}

@@

% histogram/distib for illustrated indicators for the innovation model? no time

\begin{center}
	\includegraphics[width=0.49\textwidth]{../../../../Results/InnovationMultiscale/exploration/20230623_135234_EXPLORATION/psiDiversity_MED-macroGravityDecay_color-macroInnovationDecay_facet-mesoToMacroInnovationThreshold-macroToMesoExchangeMaxUpdate_mesoCrossOverProba05.png}
	\includegraphics[width=0.49\textwidth]{../../../../Results/InnovationMultiscale/exploration/20230705-20230712_EXPLORATION/psiDiversity-macroGravityDecay_color-macroInnovationDecay_facet-mesoToMacroInnovationThreshold-macroToMesoExchangeMaxUpdate_mesoCrossOverProba05.png}
\end{center}

% not sure 100? check oms commit! -> 56547b137cb2e8c807511e71efde1c06ba42ef6d : 20 repets only!

$\rightarrow$ statistical convergence of indicators is more difficult to reach (example above for innovation diffusion: 20 vs 10,000 replications, despite ``good-looking'' empirical Sharpe ratios suggesting a low number of repetitions)

$\rightarrow$ non-linearities, fat-tail distribution, noise propagation between scales? More uncertainty-based/Bayesian validation methods?

@@

$\rightarrow$ issue of over-complication/overfitting (generic issue for simulation models \cite{raimbault2020indirect}): is the multi-scale aspect necessary?

$\rightarrow$ the stronger the emergence, the more it is needed? \cite{bedau2002downward}: for nominal emergence, can be simplified/surrogated; for weak emergence, depends on the cases? (cf Schelling model); for strong emergence difficult to be avoided because of the autonomy?

$\rightarrow$ need indicators and methods to systematically quantify emergence in multi-scalar simulated dynamics: cf. information-theory-based indicators proposed by \cite{rosas2020reconciling}

$\rightarrow$ link between strength of emergence and strength of coupling, and other types of complexities (such as \cite{pumain2003approche}'s ``interdisciplinary'' complexity), seems to be an open question, related to the nature of complexities \cite{raimbault2020relating} and complex knowledge/knowledge of the complex

@@

\sframe{Discussion: still a long way to go!}{

%\textbf{Main results:}

@@

$\rightarrow$ there seem to be methodological regularities when building strongly coupled multi-scale models - to what extent do they depend on the type of system, methodological framework, do they hold with a parametrisation on real data?

@@

$\rightarrow$ subtle compromise between model complication/complexity and the effective capture of multi-scale dynamics

@@

$\rightarrow$ application to multi-scalar policies still far, but already possible in a stylised way such as for trade-offs between SDGs (at a single scale for now \cite{raimbault2022trade})

@@

\textbf{Open repositories for presented models:}

\texttt{https://github.com/JusteRaimbault/UrbanGrowth}

\texttt{https://github.com/JusteRaimbault/InnovationMultiscale-model}
